---
title: "MPM.rmd"
author: "Renan Bonfá"
date: "15/03/2022"
output: html_document
---

Marketing predictive model developed for Waldata              
 
Developed a machine learning model that helped the company to decide wether to invest in a "Targeted Advertising" solution that predicts whether a costumer will subscribe to the premium service 

Constructed a neural network model that improved the sensitivity by 21%


# Exploratory data analysis
```{r}

library(utils)
# Setting the environment     

setwd("C:\\Users\\renan\\Downloads\\")

# Importing the dataset
df <- read.csv("Marketing.csv", sep = ";", dec = "_") 
```

```{r}
# Carregando as bibliotecas
library(dplyr)
library(ggplot2)
library(caret)
library(datasets)
library(mlbench)
library(GA)
library(doParallel)
```

###Converting numeric variables to factors
```{r}
# Converting the variables 
df$JOB <- as.factor(df$JOB)
df$MARITAL_STATUS <- as.factor(df$MARITAL_STATUS)
df$EDUCATION <- as.factor(df$EDUCATION)
df$DEFAULT <- as.factor(df$DEFAULT)
df$HOUSING <- as.factor(df$HOUSING)
df$LOAN <- as.factor(df$LOAN)
df$CONTACT <- as.factor(df$CONTACT)
df$MONTH <- as.factor(df$MONTH)# it was chosen to make it a factor, as there is no temporal information
df$DAY_OF_WEEK <- as.factor(df$DAY_OF_WEEK)
df$CAMPAIGN <- as.factor(df$CAMPAIGN)
df$POUTCOME <- as.factor(df$POUTCOME)
# For the CONS_PRICE_IDX variable, a small adjustment is needed in the values of "93_2", since they are separated by "_"
df$CONS_PRICE_IDX[df$CONS_PRICE_IDX == "93_2"] <- "93.2"
df$CONS_PRICE_IDX <- as.numeric(df$CONS_PRICE_IDX)
df$SUBSCRIBED <- as.factor(df$SUBSCRIBED)

```

### Analyzing NAs
```{r}
# Checking for NAs in the dataset
colSums(is.na(df))

# There is no presence of NAs

# Perform the dataset summary

summary(df)

# The dataset is out of balance. Only 11% of the dataset is subscription.
```

### Graphic checks, proportions and histograms
```{r}
# Checking the variables in relation to the objective variable graphically and in proportion tables.
# The histograms are also checked, when the variables are quantitative.

# Age
ggplot(data = df, aes(x = AGE, y =  SUBSCRIBED)) + geom_point()
hist(df$AGE)

## it doesn't seem to be a good predictor as there is no clear separation between age

# Job
table(df$JOB, df$SUBSCRIBED)
prop.table(table(df$JOB, df$SUBSCRIBED),2)

## It seems that being a student increases the chance of applying.

# Marital Status
table(df$MARITAL_STATUS, df$SUBSCRIBED)
prop.table(table(df$MARITAL_STATUS, df$SUBSCRIBED),2)

## Doesn't seem to be a good predictor

# Education
table(df$EDUCATION, df$SUBSCRIBED)
prop.table(table(df$EDUCATION, df$SUBSCRIBED),2)

## Doesn't seem to be a good predictor

# Default
table(df$DEFAULT, df$SUBSCRIBED)
prop.table(table(df$DEFAULT, df$SUBSCRIBED),2)

## If subscribed, then "no"

# Housing
table(df$HOUSING, df$SUBSCRIBED)
prop.table(table(df$HOUSING, df$SUBSCRIBED),2)

# Loan
table(df$LOAN, df$SUBSCRIBED)
prop.table(table(df$LOAN, df$SUBSCRIBED),2)

# Contact
table(df$CONTACT, df$SUBSCRIBED)
prop.table(table(df$CONTACT, df$SUBSCRIBED),2)

## If subscribed, then "mobile"

# Month
table(df$MONTH, df$SUBSCRIBED)
prop.table(table(df$MONTH, df$SUBSCRIBED),2)
## If may, then "no"
## If apr, then "yes"

# Day of week
table(df$DAY_OF_WEEK, df$SUBSCRIBED)
prop.table(table(df$DAY_OF_WEEK, df$SUBSCRIBED),2)

# Duration
ggplot(data = df, aes(x = DURATION, y =  SUBSCRIBED)) + geom_point()
hist(df$DURATION)
hist(log(df$DURATION))

# It may be appropriate to apply the log

# Campaign
table(df$CAMPAIGN, df$SUBSCRIBED)
prop.table(table(df$CAMPAIGN, df$SUBSCRIBED),2)

# There are some campaigns that did not work

# PDAYS
ggplot(data = df, aes(x = PDAYS, y =  SUBSCRIBED)) + geom_point()
hist(df$PDAYS)
hist(log(df$PDAYS))
# Interesting... Longer days since the last contact generated registrations
# There are enough -1

# previous
ggplot(data = df, aes(x = PREVIOUS, y = SUBSCRIBED)) + geom_point()
hist(df$PREVIOUS)
# There are enough 0

#POUTCOME
table(df$POUTCOME, df$SUBSCRIBED)
prop.table(table(df$POUTCOME, df$SUBSCRIBED),2)
## Seems to be a good predictor

# EMP_VAR_RATE
ggplot(data = df, aes(x = EMP_VAR_RATE, y = SUBSCRIBED)) + geom_point()
hist(df$EMP_VAR_RATE)

# CONS_PRICE_IDX
ggplot(data = df, aes(x = CONS_PRICE_IDX, y = SUBSCRIBED)) + geom_point()
hist(df$CONS_PRICE_IDX)

# CONS_CONF_IDX
ggplot(data = df, aes(x = CONS_PRICE_IDX, y = SUBSCRIBED)) + geom_point()
hist(df$CONS_CONF_IDX)
```

# Modeling
```{r}
# Modeling
set.seed(314)
trainIndex <- createDataPartition(df$SUBSCRIBED, p=0.7, list = FALSE)

dfTrain <- df[ trainIndex,]
dfTest  <- df[-trainIndex,]
```

### Logistic Regression
```{r}
# Logistic Regression
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE,
                   summaryFunction=twoClassSummary, classProbs = TRUE)

set.seed(314)
model_logreg <- train(SUBSCRIBED~., data = dfTrain, method = "glm", 
               metric="ROC",trControl = cv, control = list(maxit = 50))
model_logreg
summary(model_logreg)

dfPred_logreg <- predict(model_logreg, newdata=dfTest)
confusionMatrix(data=dfPred_logreg, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - Logistic Regression
```{r}
dfTest$Pred_RegLog <- dfPred_logreg
head(dfTest)
```

#### Importance - Logistic Regression
```{r}
imp_logreg <- varImp(model_logreg, useModel=FALSE, scale=FALSE)
imp_logreg
plot(imp_logreg)
```

## Decision trees

### Bagging
```{r}
## Bagging
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE, classProbs=TRUE)

set.seed(314)
model_bagging <- train(SUBSCRIBED~. , data = dfTrain, method = "treebag",trControl = cv)
model_bagging

dfPred_bagging <- predict(model_bagging ,newdata=dfTest)
confusionMatrix(data=dfPred_bagging, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - Bagging
```{r}
dfTest$Pred_bagging <- dfPred_bagging
head(dfTest)
```

#### Importance - Bagging
```{r}
imp_bagging <- varImp(model_bagging, useModel=FALSE, scale=FALSE)
imp_bagging
plot(imp_bagging)
```

### Boosting
```{r}
## Boosting
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE, classProbs=TRUE)

set.seed(314)
model_boosting <- train(SUBSCRIBED~. , data = dfTrain, method = "xgbTree",trControl = cv)
model_boosting
model_boosting$modelType

dfPred_boosting <- predict(model_boosting, newdata=dfTest)
confusionMatrix(data=dfPred_boosting, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - Boosting
```{r}
dfTest$Pred_boosting <- dfPred_boosting
head(dfTest)
```

#### Importance - Boosting
```{r}
imp_boosting <- varImp(model_boosting, useModel=FALSE, scale=FALSE)
imp_boosting
plot(imp_boosting)
```

### Random Forest
```{r}
## Random Forest
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE, classProbs=TRUE)

set.seed(314)
model_rf <- train(SUBSCRIBED~. , data = dfTrain, method = "rf",trControl = cv)
model_rf
model_rf$modelType

dfPred_rf <- predict(model_rf ,newdata=dfTest)
confusionMatrix(data=dfPred_rf, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - Random Forest
```{r}
dfTest$Pred_rf <- dfPred_rf
head(dfTest)
```

#### Importance - Random Forest
```{r}
imp_rf <- varImp(model_rf, useModel=FALSE, scale=FALSE)
imp_rf
plot(imp_rf)
```

##SVM

### SVM Linear
```{r}
##Linear
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

set.seed(314)
model_linear <- train(SUBSCRIBED~., data = dfTrain, method = "svmLinear", trControl = cv, 
                      preProcess = c("center", "scale"))

model_linear
model_linear$modelType

dfPred_linear <- predict(model_linear, newdata=dfTest)
confusionMatrix(data=dfPred_linear, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - SVM Linear
```{r}
dfTest$Pred_linear <- dfPred_linear
head(dfTest)
```

#### Importance - SVM Linear
```{r}
imp_SVM_Linear <- varImp(model_linear, useModel=FALSE, scale=FALSE)
imp_SVM_Linear
plot(imp_SVM_Linear)
```

### SVM Radial
```{r}
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

set.seed(314)
model_radial <- train(SUBSCRIBED~., data = dfTrain, method = "svmRadial", trControl = cv, 
                   preProcess = c("center", "scale"))
model_radial
model_radial$modelType

dfPred_radial <- predict(model_radial, newdata=dfTest)
confusionMatrix(data=dfPred_radial, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - SVM Radial
```{r}
dfTest$Pred_radial <- dfPred_radial
head(dfTest)
```

#### Importance - SVM Radial
```{r}
imp_SVM_Radial <- varImp(model_radial, useModel=FALSE, scale=FALSE)
imp_SVM_Radial
plot(imp_SVM_Radial)
```

### Redes Neurais
```{r}
# Neural
set.seed(314)
cv <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)

set.seed(314)
model_neural <- train(SUBSCRIBED~. , data = dfTrain, method = "nnet",
                      trControl = cv, trace = FALSE,
                      preProcess = c("center", "scale"))

model_neural
model_neural$modelType

dfPred_neural <- predict(model_neural, newdata=dfTest)
confusionMatrix(data=dfPred_neural, dfTest$SUBSCRIBED, positive = "yes")
```

#### Scoring - Neural Networks
```{r}
dfTest$Pred_neural <- dfPred_neural
head(dfTest)
```

#### Importance - Neural Networks
```{r}
imp_nnet <- varImp(model_neural, useModel=FALSE, scale=FALSE)
imp_nnet
plot(imp_nnet)
```

In an attempt to solve the problem, the techniques of Logistic Regression, Bagging, Boosting, Random Forest, SVM-Linear, SVM-Radial and Neural Networks were applied (the Genetic Algorithm was not implemented due to processing problems). All models showed satisfactory results in terms of accuracy (> 90%) and specificity (> 94%). Thus, sensitivity was chosen as the performance analysis metric of the model. In order to avoid false positives, this metric is the most appropriate, as it provides more security for the determination of targeted advertising. Therefore, the Neural Networks model was chosen. The metrics for this model were, namely, Accuracy of 90.81%, Specificity, 94.49% and Sensitivity, 61.85%.

As a future study, a balancing of the database can be applied. Note that there is little presence of positive cases (only 11% of the base is described as "yes"), which can end up biasing the models.




